---
title: "Statistical Inference Course Project - Inferential Data Analysis"
author: "Naman Gandhi"
date: "February 4, 2017"
output: pdf_document
fontsize: 10pt
geometry: margin=0.5in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this project we'll analyze the ToothGrowth data in the R datasets package.

We'll break it into 4 broad sections -  
    1) **Preparation and Data Summary**  
    2) **Exploratory Data Analysis**   
    3) **Inferential Analysis**  
    4) **Assumptions and Conclusion**  
  
  
Let's dive in!!

## Section 1: Preparation and Data Summary
This data analyses the effect of Vitamin C on Tooth Growth in Guinea Pigs
```{r}
# Loading the required libraries and dataset
library(datasets)
library(ggplot2)
data("ToothGrowth")
# Understanding the structure and dispersion of data
str(ToothGrowth)
summary(ToothGrowth)
# Converting the dose to be factor to analyze it better
ToothGrowth$dose<-as.factor(ToothGrowth$dose)
```

**Insights**  
Each of 60 animals received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).    

## Section 2: Exploratory Data Analysis

**Figure 1** Explains how the length of the tooth is affected by supplements, OJ stands for Orange Juice and VC - (Vit C) Ascorbic Acid

```{r fig.width=7, fig.height=3}
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp))+ geom_point() +
    xlab("Supplement type") +ylab("Tooth Length") +
    ggtitle("Figure 1: Tooth Length summarized by supplement")
```



**Figure 2** Explains how the length of the tooth grows as the dosage of the supplement increases, OJ stands for Orange Juice and VC - (Vit C) Ascorbic Acid
```{r fig.width=7.5, fig.height=3}
ggplot(data = ToothGrowth, aes(x = dose, y = len)) + geom_boxplot(aes(fill = dose)) + 
    facet_grid(~ supp) + geom_point(aes(colour = I("black"))) + xlab("Dose (mg)") + 
    ylab("Tooth Length") + ggtitle("Figure 2: Tooth Length vs. Dose, faceted by type of supplement")
```


## Section 3: Inferential Analysis
We'll test the significance by performing hypothesis testing in two phases -  

**3A) Testing the significance of correlation between delivery method and tooth growth  **

```{r}
t.test(len~supp,data=ToothGrowth)
```
**Insights** -  
i) The p-value of this test is 0.06  which is greater than significance level of 0.05  
ii) 95% onfidence interval contains zero  
=> *Null Hypothesis Accepted - Supplement types seems to have no impact on Tooth growth*

**3B) Testing the significance of correlation between dosage administered and tooth growth  **

```{r}
# 3B1 - Dose amounts 0.5 and 1.0
t.test(len~dose,data=subset(ToothGrowth, ToothGrowth$dose %in% c(0.5,1.0)))
```

```{r}
# 3B2 - Dose amounts 1.0 and 2.0
t.test(len~dose,data=subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,2.0)))
```

```{r}
# 3B3 - Dose amounts 0.5 and 2.0
t.test(len~dose,data=subset(ToothGrowth, ToothGrowth$dose %in% c(0.5,2.0)))
```

**Insights** -  
i) The p-value of each of the three tests (3B1, 3B2 & 3B3) test is 0,  which is less than significance level of 0.05  
ii) 95% onfidence interval does not contains zero (-ve in all cases)  
=> *Null Hypothesis Rejected - Tooth length increases with an inceasing dose*


## Section 4: Assumptions and Conclusion  
**Assumptions**  
i) The distribution of the sample means follows the Central Limit Theorem  
ii) The sample is representative of the population

**Conclusion**  
Analyzing the results from the T-tests, we can conclude that supplement delivery method has no effect on tooth growth, however increased dosages do result in increased tooth length of Guinea pigs.
