---
title: "Statistical Inference Course Project - Simulation"
author: "Naman Gandhi"
date: "February 4, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this project we'll investigate the distribution of averages of 40 exponentials in R and compare it with the Central Limit Theorem by doing a thousand simulations.  

We'll break it into 5 broad sections -  
    1) **Simulations**  
    2) **Sample Mean versus Theoretical Mean**   
    3) **Sample Variance versus Theoretical Variance**  
    4) **Understanding the Distribution**  
    5) **Conclusion**
  
  
Let's dive in!!
  
## Section 1: Simulations

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. 

```{r}
# Setting seed for reproducibility
set.seed(7815)
# Setting simulation parameters
n = 40
lambda = 0.2
simulations = 1000
# Generating 1,000 simulations
simulated_exp <- replicate(simulations, rexp(n, lambda))
# calculate columnar mean of exponentials
means_exp <- apply(simulated_exp, 2, mean)

# Visualizing the means using histogram and adding rugs to see finer density scatter
hist(means_exp, col = "orange", xlab = "Simulation Means", 
     main = "Frequency plot of Simulated means for 1,000 observations", breaks = 40)
rug(means_exp, col = "grey",lwd = 1)
```

## Section 2: Sample Mean versus Theoretical Mean
We know that the theoretical mean of an exponential distribution is 1/lambda, which implies for our distribution its 5

```{r}
(sample_mean <- mean(means_exp))        # Sample Mean
(theoretical_mean <- 1/lambda)          # Theoretical Mean
(delta_mean <- abs(sample_mean - theoretical_mean))     # Delta

```
With the delta of 0.046, we can say that The sample mean is very close to the theoretical mean

## Section 3: Sample Variance versus Theoretical Variance

``` {r}
(sample_var <- var(means_exp))          # Sample variance
(theoretical_var <- (1/lambda)^2/n)     # Theoretical variance
(delta_var <- abs(sample_var - theoretical_var))    # Delta
```
With the delta of 0.064, we can say that the sample variance is very close to the theoretical variance

## Section 4: Understanding the Distribution
The exponential distribution is approximately normal. Due to the Central Limit Theorem, the means of samples should also follow a normal distribution.
Let's see if this holds true.

```{r}
par(mfrow = c(1,2))
hist(means_exp, probability = TRUE, col = "orange", xlab = "Simulation Means", 
     main = "Histogram: Simulated means", breaks = 40)
lines(density(means_exp), lwd=3, col = "red")

qqnorm(means_exp, main="Normal Q-Q Plot", xlab="Theoretical Quantiles", ylab="Sample Quantiles")
qqline(means_exp, col="red")
```

## Section 5: Conclusion
Yes, indeed the graphs show the distribution as a close approximation of Gaussian and as we simulate for higher n sizes, the graph will tend to become more closer to the Gaussian Normal!
